DraftOne

### 2016-06-25 writing/JSSDraft/JSSDraft.Rmd
Work on JSS draft

@@

@@

@@

@@

>>>>

Simulation studies can be used to determine effective sample size or power when planning or proposing future research studies. This method can be particularly useful in cases where modeling of the study data will be particularly complex and when the strong assumptions of classical, analytical power analysis are questionable \citep{Bellan2015, Johnson2015, Burton2009}. Simulation studies can often be used to generate power analyses under a more realistic set of assumptions \cite{Burton2009}.

For example, \citet{Bellan2015} used a simulation study to determine the power and statistical validity of two potential Ebola vaccine trials, a randomized controlled trial and a stepped-wedge cluster trial. Simulation was useful in this case because the trial would be conducting in a population with clustering by geographical district and with a temporally changing hazard of Ebola infection over the trial period \citep{Bellan2015}.

\citet{Johnson2015} give an example of using simulation for a power analysis of a study of numbers of ticks on grouse chicks in a study design with multi-level clustering of the study population, both by geographic area and, within that, by brood.

\citet{Burton2009} used simulation to investigate the power of case-control studies of genetic associations for chronic diseases.

@@

A time series of simulated count outcome data with Poisson random noise can be simulated using the following model \citep{Johnson2015}:

\begin{equation}
\eta_t = \beta_0 + \sum_{m=1}^{p}\beta_{m}X_{m,t} 
\end{equation}

\begin{equation}
\eta_t = log(\lambda_t)
\end{equation}

\begin{equation}
y_i \sim Pois(\lambda_t)
\end{equation}

\noindent which is a Generalized Linear Model where $\eta_t$ is the linear predictor and a log link is used.

@@

Running a power analysis requires specifications of effect sizes, baseline counts of health outcomes, and shape of seasonal trends in exposures and outcomes. These values can be specified at realistic values based on previous research \citep{Johnson2015}; if no relevant previous research exists, in some cases a pilot study can be used to determine reasonable values to use for these inputs \citep{Johnson2015}.

@@

A number of different metrics are used to evaluate the results of simulation studies. These include:

* False positive rate / type I error rate (percent of time significant effect is detected-- i.e., null hypothesis is rejected-- when there is not a true effect) \citep{Bellan2015, Johnson2015}
* Statistical power to detect a true effect of a certain size (percent of simulations in which a true effect of a certain size is identified with a p-value of 0.05 or lower; i.e., probability of rejecting the null hypothesis when the null hypothesis is false) \citep{Bellan2015, Johnson2015}
* Precision of effect estimate (average confidence interval width over all simulations) \citep{Johnson2015}
* 95% confidence inverval coverage (percent of 95% confidence intervals from simulations that include the true effect size) \citep{Johnson2015}
* Bias in effect estimate \citep{Johnson2015}
* Margin of error [?] \citep{Johnson2015}
* Sample size required to detect a certain effect size \citep{Burton2009}
* Time to achieve required number of cases \citep{Burton2009}

@@

## Limitations

We currently do not include overdispersion when simulating health outcome counts. \citet{Johnson2015} discusses why this might be problematic and suggests some methods for incorporating overdispersion in a simulation in cases where count outcome data is expected to include overdispersion. They suggest that the following model can be used to generate the expected outcome count with overdispersion-- model the linear predictor as:

\begin{equation}
\eta_t = \beta_0 + \sum_{m=1}^{p}\beta_{m}X_{m,t} + \epsilon_t
\end{equation}

\begin{equation}
\eta_t = log(\lambda_t)
\end{equation}

\begin{equation}
y_i \sim Pois(\lambda_t)
\end{equation}

\noindent where $\epsilon_t$ represents the overdispersion and follows a normal distribution:

\begin{equation}
\epsilon_t \sim N(0, \sigma_{\epsilon}^2)
\end{equation}

In the case of no overdispersion, $\sigma_{\epsilon}^2$ equals 0. In the case of overdispersion, $\sigma_{\epsilon}^2$ will be larger than 0.
